Normalization Example using spike-in control as a standard

### 2017-08-12 examples/Example_001_ER_CTCF.Rmd
Examples finished subject to proof-reading.
Edited examples

To prepare the data for the differential binding analysis the data had to be aligned and CTCF and ER peaks called. As our CTCF control doesn't require any special aligment, data was aligned as for a standard ChIP-seq experiment. Peak calling was done with macs2 using the script shown below.

@@

As macs2 cannot differentiate between a peak arising from CTCF binding or ER binding we used two samples were squence to establish a CTCF binding consensus. One treated and one untreated. These peaks were then merged as follows.

@@

These were then used to write two DiffBind samplesheets as described in the DiffBind vignette. One using the CTCF peak list generated from the above scripts, this is the control sample sheet. The second uses the ER peaks, this is the experimetnal sample sheet. These two sample sheets are then provided as the input in the analysis below.

@@

Brundle provies a series of functions we developed for this pipeline that are used to normalizing ChIP-seq data. The usage of these functions are explained in the example that follows.

@@

Typically we would load the raw data using Diffbind as we are plan to return our normalized data as a Diffbind object.

In this example, as the BAM files which are greater in size than the github filesize limit, we have included the data as the Diffbind object these function would output and instead load that directly. If the code doesn't find the R data object it will try to regenerate if from the raw data.

@@

Finally we return the correct values into the DiffBind object. This requires a modificed version Diffbind as included in the original manuscript. It is expected that this will be supported natively by furture versions of Diffind.

>>>>

To prepare the data for the differential binding analysis the data had to be aligned and CTCF and ER peaks called. As our CTCF control doesn't require any special aligment, data was aligned as for a standard ChIP-seq experiment and followed by removal of reads from blacklisted regions. Peak calling was done with macs2 using the script shown below.

@@

As macs2 cannot differentiate between a peak arising from CTCF binding or ER binding we used samples with only CTCF antibody to establish a CTCF binding consensus. One treated and one untreated. These peaks were then merged as follows.

@@

The peak list files were then used as the basis for two DiffBind samplesheets as described in the DiffBind vignette. One using the CTCF peak list generated from the above scripts, this is the control sample sheet. The second uses the ER peaks, this is the experimetnal sample sheet. These two sample sheets are then provided as the input in the analysis below.

@@

Brundle provies a series of functions we developed for this pipeline that are used to normalize ChIP-seq data. The usage of these functions are explained in the example that follows.

@@

Typically we would load the raw data using Diffbind as we are planning to return our normalized data as a Diffbind object.

In this example, as the BAM files which are greater in size than the github filesize limit, we have included the data as the Diffbind object these function would output and instead of  loading that directly. If the code doesn't find the R data object it will try to regenerate if from the raw data.

@@

Finally we return the correct values into the DiffBind object. This requires a modified version Diffbind as included in the original manuscript. It is expected that this will be supported natively by furture versions of Diffind.

### 2017-08-12 examples/Example_003_ER_H2av.Rmd
Edited examples

This example builds on the previous two and therefore these should be read first as many details of key functions have not been repeated.

@@

With the peak files and the seperate alignments extract we are can then write samplesheets. For the normalization we need one for the drosphila aligned reads from each sample and the respective H2av peak file, and one for human reads from each sample and their repsective ER peak files. This forms the bais of the following analysis.

@@

As before to keep file size down data are provided as a RdatafFile rather than as raw BAM files.

>>>>

This example builds on the previous two and therefore they should be read first as many details of key functions have not been repeated.

@@

With the peak files and the seperate alignments extracted we are can then write samplesheets. For the normalization we need one for the drosphila aligned reads from each sample and the respective H2av peak file, and one for human reads from each sample and their repsective ER peak files. This forms the bais of the following analysis.

@@

As before to keep file size down data are provided as in Rdata format rather than as raw BAM files.
